---
title: "STAT 506, PS 4"
format: html
editor: visual
---

GitHub Repository: <https://github.com/dereklief/stats506_ps4>

## **Problem 1 - Tidyverse**

Use the **tidyverse** for this problem. In particular, use piping and **dplyr** as much as you are able. **Note**: Use of any deprecated functions will result in a point loss.

Install and load the package [**nycflights13**](https://cran.r-project.org/package=nycflights13).

```{r}
library(nycflights13)
library(tidyverse)
library(dplyr)
airports <- airports
flights <- flights
```

a.  Generate a table (which can just be a nicely printed tibble) reporting the mean and median departure delay per airport. Generate a second table (which again can be a nicely printed tibble) reporting the mean and median arrival delay per airport. Exclude any destination with under 10 flights. Do this exclusion through code, not manually.

    Additionally,

    -   Order both tables in descending mean delay.

    -   Both tables should use the airport *names* not the airport *codes*.

    -   Both tables should print all rows.

Let's work on the first table - the one for mean and median departure delay per airport, first:

```{r}
flights %>% inner_join(airports, by = c("origin"="faa")) %>% 
  group_by(name) %>% 
  summarize(delay_mean=mean(dep_delay, na.rm=TRUE),
            delay_median=median(dep_delay, na.rm = TRUE)) %>%
  arrange(-delay_mean) %>%
  print()
```

Now the second table:

```{r}
flights %>% inner_join(airports, by = c("origin"="faa")) %>% 
  group_by(dest) %>% 
  filter(n() >= 10) %>%
  group_by(name) %>% 
  summarize(arr_mean=mean(arr_delay, na.rm=TRUE),
            arr_median=median(arr_delay, na.rm = TRUE)) %>%
  arrange(-arr_mean) %>%
  print()
```

b\. How many flights did the aircraft model with the fastest average speed take? Produce a tibble with 1 row, and entries for the model, average speed (in MPH) and number of flights.

```{r}
flights %>% inner_join(planes, by = c("tailnum")) %>% 
  group_by(model) %>%
  summarize(avg_speed = mean(speed, na.rm = TRUE),
            num_flights = n()) %>%
  arrange(-avg_speed) %>%
  filter(row_number()==1)
```
